A standard for the scholarly citation of archaeological data as an incentive to data sharing

### 2018-03-11 analysis/paper/paper.Rmd
minor edits to plots while checking proofs

```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing behavious and date of publication"}
library(viridis)
library(scales) 
library(wesanderson)

@@

sharing_by_year <-
email_survey %>%
  filter(!is.na(`email sent date`)) %>%
  group_by(`publication year`, `content recieved date`) %>%
  tally() %>%
  filter(!is.na(`publication year`)) %>%
  mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  ggplot(aes(`publication year`, n)) +
    geom_col(aes(fill = `content recieved date`),
             # need devtools::install_github("tidyverse/ggplot2") for this
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)], 
                    guide = FALSE) +
  scale_x_continuous(breaks = years,
                    labels = years) +
  xlab("Publication year") +
  scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
  theme_bw() +
  coord_fixed(ratio = 0.5)

@@

```{r student, fig.cap="No relationship between student status and data sharing behavior"}
sharing_by_student <-
email_survey %>%
  filter(!is.na( `email sent date`)) %>%
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  group_by(
  `author is graduate student at time of publication?`,
  `content recieved date`) %>%
  tally() %>%
  filter(!is.na(`author is graduate student at time of publication?`)) %>%
  ggplot(aes(`author is graduate student at time of publication?`, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)],
                    guide = FALSE) +
  scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
    xlab("Author was graduate student at time of publication?") +
    theme_bw() +
  coord_fixed(ratio = 0.05)
```

```{r author-country}
sharing_by_country <-
email_survey %>%
  filter(!is.na( `email sent date`)) %>%
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  group_by(
  `author home`,
  `content recieved date`) %>%
  tally() %>%
  filter(!is.na(`author home`)) %>%
  ggplot(aes(`author home`, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)],
                    guide = FALSE) +
    scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
    theme_bw() +
  xlab("Author location at time of publication") +
  coord_fixed(ratio = 0.25)
```

@@

sharing_by_journal <-
  email_survey %>%
  filter(!is.na( `email sent date`)) %>%
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  group_by(
  journal,
  `content recieved date`) %>%
  tally() %>%
  filter(!is.na(journal)) %>%
  ggplot(aes(journal, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)],
                    guide = guide_legend(title = "Data shared?")) +
    scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = c(0.8, 0.7)) +
  xlab("Journal name") +
  coord_fixed(ratio = 0.25)
```

@@

plot_file_type_location <-
file_type_location %>%
  group_by(file, location) %>%
  tally() %>%
  ggplot(aes(file, location)) +
  geom_point(aes(size = n),
             colour = wes_palette("Rushmore")[c(2)],
             shape = 16) +
  theme_bw() +
  xlab("") +
  ylab("") +
  scale_size_continuous(range=c(7,25)) +
  geom_text(aes(label = n)) +
  theme(legend.position = "none")
```

@@

plot_data_type_location <-
data_type_location %>%
  group_by(data_type, location) %>%
  tally() %>%
  filter(n >= 2) %>%
  ggplot(aes(data_type, location)) +
  geom_point(aes(size = n),
             colour = wes_palette("Rushmore")[c(2)],
             shape = 16) +
  theme_bw() +
  xlab("") +
  ylab("") +
  scale_size_continuous(range=c(7,25)) +
  geom_text(aes(label = n), colour = "black") +
  theme(legend.position = "none")
```

@@

date_plot <-
  dates  %>%
  filter(!is.na(yyyy_date)) %>%
  filter(datacentre %in% top_data_sources[1:5]) %>% 
  ggplot(aes(yyyy_date,
             fill = datacentre
             )) +
  geom_bar() +
  # guides(fill=FALSE) +
  scale_fill_manual(values = wes_palette("Rushmore")) +
  theme_new() +
  theme(legend.position = c(0.3, 0.5)) +
  scale_x_continuous(limits = c(1950,2017), expand = c(0,0)) +
  xlab("Year data was collected") +
  ylab("Number of archaeology DOIs\nknown to DataCite") +
  geom_vline(xintercept = peak_year, colour = "red") +
  annotate("text",
           x = peak_year - 1.5,
           y = ymax,
           label = peak_year,
           size = 3) # +
  # coord_fixed(ratio = 1/1000)

@@

per_repo <-
dates  %>%
  filter(!is.na(yyyy_date)) %>%
  group_by(yyyy_date, datacentre)  %>%
  tally() %>%
  filter(n != 0,
         datacentre %in% top_data_sources[1:9]) %>%
    mutate(datacentre = str_wrap(datacentre, width = )) %>%
   ggplot(aes(yyyy_date,
              n
             )) +
  geom_line(size = 1)  +
  scale_fill_manual(values = wes_palette("Rushmore")) +
  scale_y_log10() +
  xlim(c(1990, 2017)) +
  xlab("Year DOI registered (1990-2017 only)") +
  ylab("Number of DOIs\nregistered per year") +
  theme_new() +
  facet_wrap(~ datacentre, labeller = labeller(datacentre = labels[1:9]))

@@

library(ggrepel)
plot_structured_data_files_per_file_tally <-
ggplot(structured_data_files_per_file_tally,
       aes(str_wrap(datacentre, 20),
           format)) +
      geom_point(aes(size = n),
             shape = 16,
             #alpha = 0.4,
             colour = wes_palette("Rushmore")[c(2)]) +
    geom_text_repel(aes(label = prettyNum(n, big.mark = ",")),
                  size = 2.5,
                  colour = "black",
                  segment.colour="black",
                  min.segment.length = unit(0.5, "lines"),
                  nudge_x = 0.3,
                  nudge_y = 0.5) +
  scale_size_continuous(range=c(2,25), guide = FALSE) +
    theme_bw() +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1)) +
  coord_flip() +
  xlab("Data centre") +
  ylab("File format")

@@

```{r datacite-plots}
datacentre_plot <-
  datacentre %>%
  mutate(Var1 = str_wrap(Var1, 20)) %>%
  slice(1:10) %>%
ggplot(aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip() +
  xlab("Data centre") +
  ylab("") +
  theme_new() +
  theme(axis.text.y  = element_text(size = 8))

allocator_plot <-
ggplot(allocator[1:10,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip() +
  theme_new()

creator_plot <-
ggplot(creator[1:30,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip() +
  theme_new()

language_plot <-
  language %>%
  mutate(Var2 = ifelse(grepl("en|En", Var1), "en", Var1)) %>%
  mutate(Var2 = ifelse(grepl("de|ger", Var2), "de", Var2)) %>%
  group_by(Var2) %>%
  tally(sort = TRUE) %>%
  slice(1:5) %>%
ggplot(aes(reorder(Var2, nn),
           nn)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip()  +
  theme_new() +
  xlab("Language") +
  ylab("")

contributor_plot <-
ggplot(contributor[1:20,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip()  +
  theme_new()

rights_plot <-
ggplot(rights_summary,
       aes(reorder(cc, nn),
           nn)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip()  +
  theme_new() +
  xlab("License") +
  ylab("")

publisher_plot <-
ggplot(publisher[1:10,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip() +
  theme_new()

@@

plot_datacentre_rights_table <-
datacentre_rights_table[1:11,] %>%
  ggplot(aes(cc, str_wrap(str_trunc(datacentre, side = "right", 35), 15))) +
  geom_point(aes(size = n),
             colour =  wes_palette("Rushmore")[1],
             shape = 16) +
  theme_bw() +
  xlab("License") +
  ylab("Data repository") +
  scale_size_continuous(range=c(7,25)) +
  geom_text(aes(label = prettyNum(n, big.mark = ","))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

@@

doi_lookup <- 
tribble(  ~doi,       ~doi_name,
 "10.5281",   "Zenodo",
 "10.5284",   "Medieval Archaeology",
 "10.6067",   "tDAR",
 "10.15468",  "GBIF",
 "10.14279",  "Technische Universität Berlin",
 "10.17026",  "EASY-DANS",
 "10.11141",  "Internet Archaeology",
 "10.1080",   "Various journal publishers") # OUP, CUP, Sage, Wiley, Springer

@@

library(wesanderson)
plot_citation_table <-
events_df_names %>% 
  group_by(doi_provider_name, 
           doi_consumer_name) %>% 
  tally() %>% 
  filter(!is.na(doi_provider_name)) %>% 
  filter(!is.na(doi_consumer_name)) %>% 
  ggplot(
         aes(str_wrap(str_trunc(doi_provider_name, side = "right", 35), 15),
             str_wrap(str_trunc(doi_consumer_name, side = "right", 35), 15))) +
  geom_point(aes(size = n),
             colour =  wes_palette("Rushmore")[1],
             shape = 16) +
  theme_bw() +
  xlab("Data source") +
  ylab("Cited by") +
  scale_size_continuous(range=c(5,25)) +
  geom_text(aes(label = prettyNum(n, big.mark = ","))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

@@

```{r citation-time-series}
citation_time_series_plot <-  
events_df_names %>% 
  select(occurred_at) %>% 
  mutate(occurred_at = as.Date(occurred_at)) %>% 
ggplot(aes(occurred_at)) +
  geom_histogram(fill = wes_palette("Rushmore")[4]) +
  scale_x_date(date_breaks = "6 month") +
  theme_bw(base_size = 10) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  xlab("Date of citation event") +
  ylab("Number of citations")

>>>>

```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing behavious and date of publication"}
library(scales) 
library(wesanderson)

@@

sharing_by_year <-
email_survey %>%
  filter(!is.na(`email sent date`)) %>%
  group_by(`publication year`, `content recieved date`) %>%
  tally() %>%
  filter(!is.na(`publication year`)) %>%
  mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  ggplot(aes(`publication year`, n)) +
    geom_col(aes(fill = `content recieved date`),
             # need devtools::install_github("tidyverse/ggplot2") for this
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore1")[c(1,3)], 
                    guide = FALSE) +
  scale_x_continuous(breaks = years,
                    labels = years) +
  xlab("Publication year") +
  scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
  theme_bw() +
  coord_fixed(ratio = 0.5)

@@

```{r student, fig.cap="No relationship between student status and data sharing behavior"}
sharing_by_student <-
email_survey %>%
  filter(!is.na( `email sent date`)) %>%
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  group_by(
  `author is graduate student at time of publication?`,
  `content recieved date`) %>%
  tally() %>%
  filter(!is.na(`author is graduate student at time of publication?`)) %>%
  ggplot(aes(`author is graduate student at time of publication?`, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore1")[c(1,3)],
                    guide = FALSE) +
  scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
    xlab("Author was graduate student at time of publication?") +
    theme_bw() +
  coord_fixed(ratio = 0.05)
```

```{r author-country}
sharing_by_country <-
email_survey %>%
  filter(!is.na( `email sent date`)) %>%
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  group_by(
  `author home`,
  `content recieved date`) %>%
  tally() %>%
  filter(!is.na(`author home`)) %>%
  ggplot(aes(`author home`, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore1")[c(1,3)],
                    guide = FALSE) +
    scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
    theme_bw() +
  xlab("Author location at time of publication") +
  coord_fixed(ratio = 0.25)
```

@@

sharing_by_journal <-
  email_survey %>%
  filter(!is.na( `email sent date`)) %>%
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  group_by(
  journal,
  `content recieved date`) %>%
  tally() %>%
  filter(!is.na(journal)) %>%
  ggplot(aes(journal, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore1")[c(1,3)],
                    guide = guide_legend(title = "Data shared?")) +
    scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = c(0.8, 0.7)) +
  xlab("Journal name") +
  coord_fixed(ratio = 0.25)
```

@@

plot_file_type_location <-
file_type_location %>%
  group_by(file, location) %>%
  tally() %>%
  ggplot(aes(file, location)) +
  geom_point(aes(size = n),
             colour = wes_palette("Rushmore1")[c(2)],
             shape = 16) +
  theme_bw() +
  xlab("") +
  ylab("") +
  scale_size_continuous(range=c(7,25)) +
  geom_text(aes(label = n)) +
  theme(legend.position = "none")
```

@@

plot_data_type_location <-
data_type_location %>%
  group_by(data_type, location) %>%
  tally() %>%
  filter(n >= 2) %>%
  ggplot(aes(data_type, location)) +
  geom_point(aes(size = n),
             colour = wes_palette("Rushmore1")[c(2)],
             shape = 16) +
  theme_bw() +
  xlab("") +
  ylab("") +
  scale_size_continuous(range=c(7,25)) +
  geom_text(aes(label = n), colour = "black") +
  theme(legend.position = "none")
```

@@

date_plot <-
  dates  %>%
  filter(!is.na(yyyy_date)) %>%
  filter(datacentre %in% top_data_sources[1:5]) %>% 
  ggplot(aes(yyyy_date,
             fill = datacentre
             )) +
  geom_bar() +
  # guides(fill=FALSE) +
  scale_fill_manual(values = wes_palette("Rushmore1")) +
  theme_new() +
  theme(legend.position = c(0.3, 0.5)) +
  scale_x_continuous(limits = c(1950,2017), expand = c(0,0)) +
  xlab("Year data was collected") +
  ylab("Number of archaeology DOIs\nknown to DataCite") +
  geom_vline(xintercept = peak_year, colour = "red") +
  annotate("text",
           x = peak_year - 1.5,
           y = ymax,
           label = peak_year,
           size = 3) # +
  # coord_fixed(ratio = 1/1000)

@@

per_repo <-
dates  %>%
  filter(!is.na(yyyy_date)) %>%
  group_by(yyyy_date, datacentre)  %>%
  tally() %>%
  filter(n != 0,
         datacentre %in% top_data_sources[1:9])  %>%
   ggplot(aes(yyyy_date,
              n,
             group = datacentre)) +
  geom_line(size = 1.5)  +
  scale_fill_manual(values = wes_palette("Rushmore1")) +
  scale_y_log10() +
  xlim(c(1990, 2017)) +
  xlab("Year DOI registered (1990-2017 only)") +
  ylab("Number of DOIs\nregistered per year") +
  theme_new() +
  facet_wrap( ~datacentre, labeller = labeller(datacentre = labels[1:9]))

@@

library(ggrepel)
plot_structured_data_files_per_file_tally <-
ggplot(structured_data_files_per_file_tally,
       aes(str_wrap(datacentre, 20),
           format)) +
      geom_point(aes(size = n),
             shape = 16,
             #alpha = 0.4,
             colour = wes_palette("Rushmore1")[c(2)]) +
    geom_text_repel(aes(label = prettyNum(n, big.mark = ",")),
                  size = 2.5,
                  colour = "black",
                  segment.colour="black",
                  min.segment.length = unit(0.5, "lines"),
                  nudge_x = 0.3,
                  nudge_y = 0.5) +
  scale_size_continuous(range=c(2,25), guide = FALSE) +
    theme_bw() +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1)) +
  coord_flip() +
  xlab("Data center") +
  ylab("File format")

@@

```{r datacite-plots}
datacentre_plot <-
  datacentre %>%
  mutate(Var1 = str_wrap(Var1, 20)) %>%
  slice(1:10) %>%
ggplot(aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore1")[4]) +
  coord_flip() +
  xlab("Data centre") +
  ylab("") +
  theme_new() +
  theme(axis.text.y  = element_text(size = 8))

allocator_plot <-
ggplot(allocator[1:10,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore1")[4]) +
  coord_flip() +
  theme_new()

creator_plot <-
ggplot(creator[1:30,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore1")[4]) +
  coord_flip() +
  theme_new()

language_plot <-
  language %>%
  mutate(Var2 = ifelse(grepl("en|En", Var1), "en", Var1)) %>%
  mutate(Var2 = ifelse(grepl("de|ger", Var2), "de", Var2)) %>%
  group_by(Var2) %>%
  tally(sort = TRUE) %>%
  slice(1:5) %>%
ggplot(aes(reorder(Var2, nn),
           nn)) +
  geom_col(fill = wes_palette("Rushmore1")[4]) +
  coord_flip()  +
  theme_new() +
  xlab("Language") +
  ylab("")

contributor_plot <-
ggplot(contributor[1:20,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore1")[4]) +
  coord_flip()  +
  theme_new()

rights_plot <-
ggplot(rights_summary,
       aes(reorder(cc, nn),
           nn)) +
  geom_col(fill = wes_palette("Rushmore1")[4]) +
  coord_flip()  +
  theme_new() +
  xlab("License") +
  ylab("")

publisher_plot <-
ggplot(publisher[1:10,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore1")[4]) +
  coord_flip() +
  theme_new()

@@

plot_datacentre_rights_table <-
datacentre_rights_table[1:11,] %>%
  ggplot(aes(cc, str_wrap(str_trunc(datacentre, side = "right", 35), 15))) +
  geom_point(aes(size = n),
             colour =  wes_palette("Rushmore1")[1],
             shape = 16) +
  theme_bw() +
  xlab("License") +
  ylab("Data repository") +
  scale_size_continuous(range=c(7,25)) +
  geom_text(aes(label = prettyNum(n, big.mark = ","))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

@@

doi_lookup <- 
tribble(  ~doi,       ~doi_name,
 "10.5281",   "Zenodo",
 "10.5284",   "Medieval Archaeology",
 "10.6067",   "tDAR",
 "10.15468",  "GBIF",
 "10.14279",  "Technische Universität Berlin",
 "10.17026",  "DANS-EASY",
 "10.11141",  "Internet Archaeology",
 "10.1080",   "Various journal publishers") # OUP, CUP, Sage, Wiley, Springer

@@

library(wesanderson)
plot_citation_table <-
events_df_names %>% 
  group_by(doi_provider_name, 
           doi_consumer_name) %>% 
  tally() %>% 
  filter(!is.na(doi_provider_name)) %>% 
  filter(!is.na(doi_consumer_name)) %>% 
  ggplot(
         aes(str_wrap(str_trunc(doi_provider_name, side = "right", 35), 15),
             str_wrap(str_trunc(doi_consumer_name, side = "right", 35), 15))) +
  geom_point(aes(size = n),
             colour =  wes_palette("Rushmore1")[1],
             shape = 16) +
  theme_bw() +
  xlab("Data source") +
  ylab("Cited by") +
  scale_size_continuous(range=c(5,25)) +
  geom_text(aes(label = prettyNum(n, big.mark = ","))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

@@

```{r citation-time-series}
citation_time_series_plot <-  
events_df_names %>% 
  select(occurred_at) %>% 
  mutate(occurred_at = as.Date(occurred_at)) %>% 
ggplot(aes(occurred_at)) +
  geom_histogram(fill = wes_palette("Rushmore1")[4]) +
  scale_x_date(date_breaks = "6 month") +
  theme_bw(base_size = 10) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  xlab("Date of citation event") +
  ylab("Number of citations")
